---
title: "Week 3 Exercises"
author: "J.Blondeau"
date: "2023-09-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All the following question use the sttst_all_sites.xlsx file in the /data directory.

All answers should be made in the .Rmd file, showing the code and your answer and everything pushed back up to your GitHub repo.

triple stars renders a thin line across the page

***
1. Import sttstj_all_sites and inspect the data set.  What is the data structure and the data type of each field?
#Imported. Make sure to change "sheet" in the bottom left to the name of the excel tab that you want to work with
subset(yr_site)

```{r}
 #this is a code chunk, anything inside the back tics is like being in the R console.



```
***

install.packages("tidyverse")
library(tidyverse)

#2. How many samples are in the data set?  
#2505

#3. How many unique habitat types are there?  What are the names of each habitat?
unique(sttstj_all_sites$hab)
#6  unique sites
# "PVMT" "AGRF" "SCR"  "PTRF" "BDRK" "HARD"

#4. How many sites are there in the "PVMT" habitat type?
subset(sttstj_all_sites, subset= hab== 'PVMT')
#10 sites

#5. How many sites are there in the "PVMT" or "AGRF" habitat types?
#10

sttstj_all_sites %>%
filter(hab == 'PVMT' | hab== 'AGRF')

#This is another way to do the same thing
sttstj_all_sites %>%
filter(hab %in% c('PVMT', 'AGRF'))

#If you make a vector of the sites you can find out how many sites by doing the below
PA<- c('PVMT', 'AGRF')

sttstj_all_sites %>%
filter(hab %in% PA)


6. How many sites are there in all the habitat types EXCEPT "PVMT"?

NoPVMT<-sttstj_all_sites%>%filter(hab!='PVMT')
#1934
#Made a value called NoPVMT

#7. Show the first 5 records (rows) of...
#    - all sites sampled in "AGRF" or "SCR"
#    - only show the "yr_site" and "hab" columns

AoS<-sttstj_all_sites%>%filter(hab=='AGRF'| hab == 'SCR')%>% select(yr_site, hab) 
AoS[1:5,]    
    
<!-- extra line break -->
<br>

#8. What is the northern most site sampled in the "BDRK" habitat?
BR<-sttstj_all_sites%>%filter(hab=='BDRK')%>%arrange(desc(lat))
BR[1,]
#2019_5450
#arrange does smallest to biggest so we need to tell it to desc so it does the opposite


#9. What is the western most site sampled in either "AGRF" or "PTRF" habitat?
AoP<-sttstj_all_sites%>%filter(hab=='AGRF' | hab== 'PTRF')%>%arrange(lon)
AoP[1,]
#2015_9758


#10. How many years were surveyed in the data set?

split<-sttstj_all_sites %>% separate(yr_site, c('year','site'))
View(split)
unique(split$year)
length(unique(split$year)) #this will count the returns for you
#16

#11. What is the southern most site sample in the "PVMT" habitat in 2004?

P04<-split %>% filter(hab=='PVMT'& year=='2004')%>%arrange(lat)
View(P04)
P04[1,]
#1932
